---
title: "Exercises for the meeting of May 10, 2019"
author: "Violeta Sosa"
date: "May 09, 2019"
output:
  html_document:
    toc: true
    toc_float: true
    theme: simplex
    highlight: tango
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(sf)
library(tidyverse)
library(units)
library(lwgeom)


```

## Exercises for the meeting of May 10, 2019

Exercises related to simple features in R (1). Try to upload your answer to the LearnWeb assigned as a .zip file (not a .rar file) containing two files: (i) an R markdown file and (ii) an html file that resulted from knitting the R markdown file. (If this is problematic, use pdf or something else.)

##Exercise 1
Look for a shapefile; download it; import it with `sf::st_read`, and plot it. What does it show?

```{r warning=FALSE}
shapeName <- st_read('depto/depto.shp')
shapeName
plot(shapeName, main="Colombia shapefile")
```


_The plot shows a map per each one of the attributes defined in the shapefile's attribute table._

##Exercise 2
Try plotting one of the attribute variables using a different color scheme

```{r warning=FALSE}
shapeName%>%select(AREA) %>%plot(graticule = TRUE, axes = TRUE)
shapeName%>%select(AREA) %>%plot(graticule = TRUE, axes = TRUE, col = heat.colors(10))
shapeName%>%select(DPTO) %>%plot(graticule = TRUE, axes = TRUE, col = topo.colors(50))

```

##Exercise 3
What is the class of the object plotted? Plot only the geometry using `st geometry`
```{r warning=FALSE}
class(shapeName)
plot(st_geometry(shapeName), main="Colombia Geometry")
```

##Exercise 4
Which layers does the GeoPackage http://www.geopackage.org/data/sample1_2.gpkg have? Import them both in R, using `sf`. Call the object with feature geometries `x1`.
```{r warning=FALSE}
st_layers("sample1_2.gpkg")

dataset <- ("sample1_2.gpkg")

counties<-read_sf(dataset, "counties")
countiestbl<-read_sf(dataset, "countiestbl")

x1<-counties
x1
x1%>%select(AREA) %>%plot(graticule = TRUE, axes = TRUE, main="Colombia (Only) Geometry Plot")
```

##Exercise 5
Create an `sf` object with a point geometry at latitude `36.21` and longitude `-81.19`

```{r warning=FALSE}
(sfc = st_sfc(st_point(c(-81.19,36.21)), crs = 4269))
```

##Exercise 6
Query the features of x1 at this point: plot the resulting geometry, and show the attributes associated with this geometry.

```{r warning=FALSE}
geometry=st_intersects(sfc,x1)
geometry

plot(st_geometry(x1[unlist(geometry),]),expandBB = c(2,2,2,2), col=rgb(red = 1, green = 0, blue = 0, alpha = 0.5), main="Resulting Geometries")
plot(st_geometry(x1), add=TRUE)
plot(sfc, expandBB = c(10, 10, 10, 10), col = "red", lwd = 5, add=TRUE)
x1[unlist(geometry),]

```

##Exercise 7
For the `nc` dataset that ships with package `sf`, compute the area of county Columbus

```{r warning=FALSE}
(file <- system.file("gpkg/nc.gpkg", package="sf"))
(file %>% read_sf()  -> nc)

nc%>%select(NAME)%>%filter(NAME=="Columbus") %>%plot(graticule = TRUE, axes = TRUE, col=rgb(red = 1, green = 0, blue = 0, alpha = 0.5), ADD=TRUE, main="Columbus County")
columbus = nc %>% filter(NAME == "Columbus")
st_area(columbus)

```

##Exercise 8
Compute also the same area after transforming the geometry to `EPSG 2264`, and compare the value with that obtained from the unprojected data; express the difference in a percentage.

```{r}
nc_newgeo<-nc
(nc_newgeo %>% st_transform(2264) -> nc_newgeo.2264)
columbus_new = nc_newgeo.2264 %>% filter(NAME == "Columbus")
(area4267=st_geod_area(columbus))
(area2264=(st_area(columbus_new) %>% set_units(m^2)))

#New area in percentage
((area2264/area4267)*100)

```

##Exercise 9
Create a plot with the `nc` states outlines that has all states that are partly within 50 km from Columbus filled with color green.

```{r}
columbusarea = nc[nc$NAME == "Columbus", ]
(selection= st_is_within_distance(columbusarea,nc, dist = 50))
plot(st_geometry(nc), main="States partly within 50km from Columbus County")
plot(st_geometry(nc[unlist(selection),]), add = TRUE, col = 'green')

```

##Exercise 10
Create a plot with the `nc` states outlines that has all states that are entirely within 100 km from Columbus filled with color red.
```{r}
(columbusarea = nc[nc$NAME == "Columbus", ])
plot(st_geometry(nc),  main="States entirely within 100km from Columbus County")

#Columbus transform for buffer distance calculation 
nc_tr = st_transform(columbusarea, 3857)
plot(st_transform(st_buffer(nc_tr, 100000),4267),col=rgb(red = 1, green = 0, blue = 0, alpha = 0.5), border = rgb(red = 1, green = 0, blue = 0, alpha = 0.5), add = TRUE)
otro=st_transform(st_buffer(nc_tr, 100000),4267)
contain_true=st_contains(otro,nc)
plot(st_geometry(nc[unlist(contain_true),]), add = TRUE, col = 'red')

```
