---
title: "Spatial data science with R"
author: "Violeta Sosa"
date: "15 de mayo de 2019"
output:
  html_document:
    toc: true
    toc_float: true
    theme: simplex
    highlight: tango
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises for the meeting of May 17, 2019

Hand in: zip file with .Rmd and .html file, with answers to the exercises of the chapter: https://keen-swartz-3146c4.netlify.com/featureattributes.html as well as those in the "further exercises" html for today.

## Exercise 1

Add a variable to the `nc` dataset by `nc$State = "North Carolina"`. Which value should you attach to this variable for the attribute-geometry relationship (agr)?
```{r warning=FALSE}
library(sf)
library(tidyverse)

nc <- system.file("gpkg/nc.gpkg", package="sf") %>%
  read_sf() %>%
  st_transform(32119)
nc2 <- nc %>% select(BIR74, SID74, NAME) %>%
  st_set_agr(c(BIR74 = "aggregate", SID74 = "aggregate", NAME = "identity"))

nc2%>%st_agr()

nc
nc$State = "North Carolina"
nc
nc1 <- nc %>% select(NAME,State) %>%
  st_set_agr(c(State= "constant"))

nc1 %>%
  st_agr()
```

_We should attach a constant agr because this value is valid for all the geometries_

## Exercise 2

Create a new sf object from the geometry obtained by `st_union(nc)`, and assign `"North Carolina"` to the variable `State`. Which `agr` can you now assign to this attribute variable?
```{r warning=FALSE}
(state_geom=st_sf(st_union(nc)))
plot(state_geom)
state_geom$State="North Carolina"

st2<-state_geom%>%select(State)%>%
  st_set_agr(c(State="identity"))

st2%>%
  st_agr()

st2
```
_This variable State is now a `identity variable` because it represents the whole geometry_

## Exercise 3

Use `st_area` to add a variable with name `area` to `nc`. Compare the `area` and `AREA` variables in the `nc` dataset. What are the units of `AREA`? Are the two linearly related? If there are discrepancies, what could be the cause?

```{r warning=FALSE}
nc$area=st_area(nc)
nc%>%select(NAME, AREA,area)
```

_`AREA` units are in arc degrees corresponding to 100 000 meters in the metric system. They are related by `AREA=area m^2 * 100000 m^2`. Being more strict if `L` is an arc degree and `r` is the Earth radius in meters we have_

$$L=\frac{2πr}{360}$$

_Calculating `L` and by replace `r`_ 

$$L=\frac{2π*6378000}{360}$$

_we have that an arc degree is `113317.099 meters`. The discrepancies are because of the projections system. AS the Earth shape behaves as a geoide, there are differences in the areas projected per CRS._

## Exercise 4

Is the `area` variable intensive or extensive? Is its `agr` equal to `constant`, `identity` or `aggregate`?

```{r warning=FALSE}
nc%>%select(NAME, AREA,area)%>%
  st_agr()

```
_The `area` variable is extensive because of its property of `change` if it is splitted for example. It is mutable. It is `agr` is equal to `aggregate` according to its nature of being a "summary value over the geometry" [1]_ 

## Exercise 5

Find the name of the county that contains POINT(-78.34046 35.017)

```{r warning=FALSE}
#Create the point
look<-st_sfc(st_point(c(-78.34046,35.017)),crs=4326)
plot(st_geometry(st_transform(nc,4326)), col = terrain.colors(50,alpha=0.5),main="Counties and Point")
plot(look,col="red",pch = 19, add=TRUE)
#Query the intersection
container=(st_intersects(look,st_transform(nc,4326)))
states<-nc[unlist(container),]
states$NAME
#Ploting the result
plot(st_geometry(states),col="orange",main="States that intersects POINT(-78.34046 35.017)")
```

## Exercise 6

Find the names of all counties with boundaries that touch county `Sampson`.

```{r warning=FALSE}
sampson= nc[nc$NAME == "Sampson", ]
sampson_to=nc[unlist(st_touches(sampson,nc)),]
#List of names
sampson_to$NAME
plot(st_geometry(sampson_to),col="orange",main="Counties with boundaries that touch county `Sampson`.")
plot(st_geometry(sampson), col="red",add=TRUE)

```

## Exercise 7

List the names of all counties that are less than 50 km away from county `Sampson`.

```{r warning=FALSE}
sampson_close=nc[unlist(st_is_within_distance(sampson,nc,dist=50000)),]
sampson_close$NAME
```

For the following exercises, we will work with naturalearth:

```{r warning=FALSE}
library(rnaturalearth)
e = ne_countries(returnclass = "sf")["pop_est"]
```


## Exercise 8

Plot the variable pop_est. Which projection is used?

```{r warning=FALSE}
e
plot(e)  
```

_The projection used is WGS84_

## Exercise 9

Plot the variable converted to +proj=ortho, see here. What happened to Antarctica?

```{r warning=FALSE}
e2<-st_transform(e,st_crs("+proj=ortho"))
plot(e2)
```

_Antarctica is no longer visible in the plot_

## Exercise 10

Plot the same data with the ortho projection, adding `lat_0=-90`, `lat_0=90`, `lon_0=90`, `lon_0=-90` for four different views. Try to explain what happens.

```{r warning=FALSE}
plot(st_transform(e,st_crs("+proj=ortho +lat_0=-90")),col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - South Pole view")
plot(st_transform(e,st_crs("+proj=ortho +lat_0=90")),col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - North Pole view")
plot(st_transform(e,st_crs("+proj=ortho +long_0=90")),col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - Side Pole view")
plot(st_transform(e,st_crs("+proj=ortho +long_0=-90")),col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - Side Pole view")

```

## Exercise 11

Create a polygon of the northern hemisphere. Use that to get the intersection with e, and plot that with `lat_0=90`.

```{r warning=FALSE}
#Get North Hemisphere
hnorth<-st_transform(e,st_crs("+proj=ortho +lat_0=90"))
plot(hnorth,col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - North Pole view")
#Create a unique geometry
hnorthclean<-st_union(st_geometry(e[unlist((!is.na(st_dimension(hnorth)))),]))
plot(hnorthclean)
#Plot the new geometry from pole north view
hnorthcleanview<-st_transform(hnorthclean,st_crs("+proj=ortho +lat_0=90"))
plot(hnorthcleanview,col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - North Pole view")

#Intersection with e
(result<-st_intersects(hnorthclean,e,sparse=FALSE))
geom_in<-e[unlist(result),]

result2<-st_transform(geom_in,st_crs("+proj=ortho +lat_0=90"))
#Intersection plotting
plot(result2,col = terrain.colors(50,alpha=0.5),main="Natural Earth Dataset - North Pole view result")

```

## Exercise 12

For the land mass between latitudes 45 and 55 and longitudes 0 and 10, estimate the total population as well as the average population.

```{r warning=FALSE}
landmass<-st_crop(e, c(xmin=0, xmax=10, ymin=45, ymax=55))
plot(landmass,col = terrain.colors(10,alpha=0.5),main="Land Mass crop for Natural Earth Dataset")
(total_pop=sum(landmass$pop_est))
(avg_pop=mean(landmass$pop_est))
```

## References
_[1] Spatial Data Science_ https://keen-swartz-3146c4.netlify.com/

_[2] R Documentation_ https://www.rdocumentation.org

_[3] sf Github_ https://r-spatial.github.io/sf/articles/sf1.html
