---
title: "Spatial data science with R"
author: "Violeta Sosa"
date: "June 6, 2019"
output:
  html_document:
    toc: true
    toc_float: true
    theme: simplex
    highlight: tango
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assigment 7

Two data sets are found here:
  
http://pebesma.staff.ifgi.de/pm10/3_2a.txt 
https://github.com/r-spatial/stars/blob/master/inst/tif/L7_ETMs.tif

## Exercise 1
Read the dataset `3_2a.txt` in, and try to create the showed time series plot

```{r warning=FALSE}
library(readr)
library(ggplot2)
datat=data.frame(read_csv("3_2a.txt"))

plot(type="l",main="Time Series Plotting", xlab="Time",ylab="pm10",strptime(datat$Time, format="%H:%M"), datat$pm10)

ggplot(datat, aes(datat$Time,datat$pm10)) + geom_line() +labs(x = "Time",y="pm10", title="Time Series Plotting")

```

## Exercise 2
For the following plot use the same dataset; you may want to use segments to color line segments according to their `pm10` value (where `pm10` is linearly mapped into `sf::sf.colors(10)`)


```{r warning=FALSE}
(colores=rev(sf::sf.colors(10)))
(rangeCol=100*seq(0,max(datat$pm10),length.out=10))

plot(datat$Long,datat$Lat, col=colores,main="Segments Plot", xlab="Lon",ylab="Lat")

s <- seq(length(datat$pm10)-1)
s <- s[-length(s)]

segments(datat$Long[s], datat$Lat[s], datat$Long[s+2], datat$Lat[s+2],rep(colores, rangeCol), lwd = 5)
```

## Exercise 3
Try to reproduce the following plot, where pixel values were mapped into `grey((9:0)/10)` for the first band of `L7_ETMs.tif`

```{r warning=FALSE}
library(stars)
library(sf)

tif = system.file("tif/L7_ETMs.tif", package = "stars")
x = read_stars(tif)
colorsGrey=(grey((9:0)/10))
colorsGrey
plot(x[,,,1], col=colorsGrey, breaks=c(seq(47,255,length.out = 11)))
```


## References

[[1] R Stars Library Documentation](https://r-spatial.github.io/stars/reference)

[[2] R Documentation](https://www.rdocumentation.org)

[[3] R Bloggers](https://www.r-bloggers.com/using-line-segments-to-compare-values-in-r/)
